---
title: "Waterman Workspaces: Customer Retention Through Case Management"
author: "Bilal Raja"
format: 
  revealjs: 
    theme: default
    fontsize: 24pt
    toc: false
    slide-number: true
    transition: fade
    fig-align: center
    margin: 0.1
    slide-layout: wide
    controls: true
    progress: true
    hash: true
    scrollable: true
    logo: images/logo.png
    title-slide-attributes: 
      data-background-color: "#005E5E"  # Dark green background
      data-background-opacity: "1"
      data-color: "white"  # White text
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE,
  fig.align = "center",
  out.width = "80%"
)

library(tidyverse)
library(readxl)
library(visdat)
library(broom)
library(kableExtra)
library(DT)

wmcases <- read_xlsx("../data/CRM Cases 1-08-2024 5-17-55 PM.xlsx")
memberships <- read_xlsx("../data/All Products 2023 to now.xlsx")

cases_clean <- wmcases |>
  filter(!grepl("Product delivery", Subject, ignore.case = TRUE) &
           !grepl("ETC waiver", Subject, ignore.case = TRUE) &
           !grepl("Financial request", Subject, ignore.case = TRUE) &
           !grepl("Purchase orders", Subject, ignore.case = TRUE) &
           !grepl("Rent/Membership assistance", Subject, ignore.case = TRUE) &
           !grepl("All retired ones", Subject, ignore.case = TRUE) &
           !grepl("\\(RETIRED\\)", Subject, ignore.case = TRUE)) |>
  filter(!is.na(`Account Number (Customer) (Account)`)) |>
  rename(`Account Number` = `Account Number (Customer) (Account)`) |>
  select(-`(Do Not Modify) Case`, -`(Do Not Modify) Row Checksum`, -`(Do Not Modify) Modified On`)

member_clean <- memberships |>
  filter(!grepl("Quote Duplicate", `Status Reason`, ignore.case = TRUE)) |>
  filter(!grepl("Casual Hire", `Accounting Code`, ignore.case = TRUE)) |>
  filter(!is.na(`Account Number (Lessee) (Account)`)) |>
  rename(`Account Number` = `Account Number (Lessee) (Account)`)

cases_joined <- left_join(cases_clean, member_clean, by = "Account Number")

cases_joined <- cases_joined |>
  select(`Account Number`, `Customer`, `Case Number`, `Case Title`, `Case Age`, `Case Age (Days)`, 
         `Is Escalated`, everything(), -Priority, -`Modified On`, -Satisfaction, -`Sentiment Value`, 
         -`Service Level`, -SLA, -Severity, -Status.y, -`Product Name`, -Location, -`Accounting Code`, 
         -`Status Reason.x`, -`Industry (Lessee) (Account)`, -Lessee) |>
  mutate(
    Site = as.factor(Site),
    `Product Category` = as.factor(`Product Category`)) |>
  filter(!grepl("TestPayment", `Account Number`, ignore.case = TRUE)) |>
  filter(!grepl("Waterman Workspaces", `Customer`, ignore.case = TRUE))

churn_statuses <- c("Inactive - Customer Cancelled", "Off-boarding", "Inactive")

cases_filtered <- cases_joined |>
  filter(`Status Reason.y` %in% churn_statuses | is.na(`Status Reason.y`)) |>
  mutate(`Is Escalated` = ifelse(`Is Escalated` == "Yes", 1, 0))


cases_count <- cases_filtered |>
  group_by(`Account Number`) |>
  summarise(num_cases = n())

avg_case_age <- cases_filtered |>
  group_by(`Account Number`) |>
  summarise(avg_case_age = median(`Case Age (Days)`, na.rm = TRUE))

customer_features <- cases_count |>
  inner_join(avg_case_age, by = "Account Number")

customer_features <- cases_filtered |>
  group_by(`Account Number`) |>
  summarise(
    num_cases = n(),
    avg_case_age = median(`Case Age (Days)`, na.rm = TRUE),
    churned = as.integer(any(`Status Reason.y` %in% churn_statuses)),
    product_category = first(`Product Category`),
    site = first(`Site`),
    is_escalated = max(`Is Escalated`, na.rm = TRUE)
  )

customer_features <- customer_features |>
  mutate(
    product_category = as.factor(product_category),
    site = as.factor(site)
  )

# Outlier detection for average case age
Q1_age <- quantile(customer_features$avg_case_age, 0.25, na.rm = TRUE)
Q3_age <- quantile(customer_features$avg_case_age, 0.75, na.rm = TRUE)
IQR_age <- Q3_age - Q1_age

lower_bound_age <- Q1_age - 1.5 * IQR_age
upper_bound_age <- Q3_age + 1.5 * IQR_age

customer_features_clean <- customer_features |>
  filter(avg_case_age >= lower_bound_age & avg_case_age <= upper_bound_age)

# Outlier detection for number of cases
Q1_num <- quantile(customer_features_clean$num_cases, 0.25, na.rm = TRUE)
Q3_num <- quantile(customer_features_clean$num_cases, 0.75, na.rm = TRUE)
IQR_num <- Q3_num - Q1_num

lower_bound_num <- Q1_num - 1.5 * IQR_num
upper_bound_num <- Q3_num + 1.5 * IQR_num

customer_features_clean <- customer_features_clean |>
  filter(num_cases >= lower_bound_num & num_cases <= upper_bound_num)

log_model <- glm(churned ~ num_cases + avg_case_age + is_escalated + site, data = customer_features_clean, family = binomial)

customer_features_clean <- customer_features_clean |>
  mutate(
    num_cases_escalated = num_cases * is_escalated,
    case_age_escalated = avg_case_age * is_escalated,
    site_escalated = interaction(site, is_escalated)
  )

int_model <- glm(churned ~ num_cases + avg_case_age + is_escalated + site + num_cases_escalated + case_age_escalated, data = customer_features_clean, family = binomial)

```

# Introduction {background-color="#005E5E"}

## About Me

::: {.incremental}

- Where am I from?
  
  - Born and raised in Jhelum, Pakistan.

- What am I studying?

  - Final Year Master of Business Analytics from Monash University.

  - Undergraduate from Monash Malaysia.
    
- I LOVE football.

:::

## Background

::: {.incremental}
- **Analysis Overview**:  
  Conducted a comprehensive analysis of customer CRM cases to identify trends and characteristics associated with customer churn.

- **Data Collection**:  
  Utilized CRM data to assess various features, including case frequency, average case age, escalation status, and location.

- **Methodology**:  
  Employed logistic regression modeling to evaluate the impact of these features on customer retention and churn likelihood.
:::

## Why This Analysis Matters

::: {.incremental}
- **Customer Retention**:  
  Retaining existing customers is more cost-effective than acquiring new ones, making it essential to identify churn risk factors.

- **Targeted Strategies**:  
  Insights from the analysis enable the development of tailored strategies to address specific issues leading to churn.

- **Data-Driven Culture**:  
  Leveraging data fosters a culture of informed decision-making within the organization, enhancing overall business performance.

- **Long-Term Success**:  
  Understanding and mitigating churn is crucial for sustaining revenue growth and building lasting customer relationships.
:::




# Methodology {background-color="#005E5E"}

## Data Preparation

::: {.incremental}
- **Analysis Overview**:  
  Prepared CRM data to identify key factors influencing customer churn.

- **Data Sources**:  
  - Integrated CRM case data.
  - Included membership information to analyze account-related factors.

- **Data Cleaning**:  
  - Removed irrelevant cases and ensured consistency in key variables to enhance data integrity for analysis.
  - Joined the CRM Cases dataset with the Membership dataset
:::

## Key Variables

::: {.incremental}
- **Churn Definition**:  
  Established churn based on specific inactive statuses, such as "Inactive - Customer Cancelled," to accurately assess customer retention.

- **Customer Metrics**:  
  Developed key features including:
  - **Number of Cases**: Frequency of customer interactions.
  - **Average Case Age**: Duration of cases to gauge responsiveness.
  - **Escalation Status**: Indicator of case urgency and potential customer dissatisfaction.
  - **Site**: Location of Waterman Offices
:::

# Exploratory Data Analysis (EDA) {background-color="#005E5E"}

## Distribution of Cases {.smaller}

Understanding the distribution of cases by their number and average resolution age.

:::: panel-tabset
## Number of Cases {.smaller}

:::: {.columns}

::: {.column width="80%"}

```{r}
ggplot(customer_features_clean, aes(x = num_cases, fill = as.factor(churned))) +
    geom_histogram(position = "dodge", binwidth = 1) +  # Use dodge to separate bars by churn status
  labs(
    title = "Distribution of Customers by Number of Cases",
    x = "Number of Cases", 
    y = "Number of Customers",
    fill = "Customer Status"
  ) +
  scale_fill_manual(values = c("0" = "steelblue", "1" = "#F08080"), 
                    labels = c("0" = "Did Not Churn", "1" = "Churned")) +  # Update labels
  theme_minimal() +
  theme(legend.position = "right")

```

:::

::: {.column width="20%"}
- Churn Rates decrease with increasing number of cases.
- Varying proportionality between churned and not churned.

:::

::::

## Average Case Age {.smaller}

:::: {.columns}

::: {.column width="70%"}
```{r}
ggplot(customer_features_clean, aes(x = avg_case_age, fill = as.factor(churned))) +
   geom_histogram(position = "dodge", binwidth = 1) + 
  labs(
    title = "Distribution of Customers by Average Case Age",
    x = "Average Case Age (Days)", 
    y = "Number of Customers",
    fill = "Customer Status"
  ) +
  scale_fill_manual(values = c("0" = "steelblue", "1" = "#F08080"), 
                    labels = c("0" = "Did Not Churn", "1" = "Churned")) +
  theme_minimal() +
  theme(legend.position = "right")


```
:::

::: {.column width="30%"}
  - Churn Rates decrease with increasing average age of case resolutions.
  - Varying proportionality between churned and not churned.
  
- Could the varying proportionality be due to chance? Worth some digging.
:::

::::

::::

# Results {background-color="#005E5E"}


## Model Key Findings: What May Impact Churn?

::: {.incremental}

- **Key Predictors**:  
  Significant predictor: **Is Escalated** (p < 0.001).

- **Non-significant Factors**:  
  - **Num Cases**: p = 0.458  
  - **Avg Case Age**: p = 0.651  
  - **Site**: No significant differences across sites.

- **Interpretation**:  
  - Escalation status is a strong indicator of churn likelihood; other factors require further exploration.
  - Intercept: When all predictors are at zero, likelihood of churning is very low.

:::

# Recommendations and Limitations {background-color="#005E5E"}

## Future Directions
::: {.incremental}
- **Proactive Engagement**: Reach out to customers with a high number of cases or long resolution times.
- **Monitoring Escalations**: Develop protocols for cases marked as escalated, ensuring timely follow-up.
- **Data Completeness**: Data extracted from CRM incomplete, require further key variables.
- **Short Time Frame**: Analysis is based on a limited period; longer-term trends could differ.
:::

# Q/A {background-color="#005E5E"}

